Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-05 Manuscript/manuscript.tex
Update binding affinity section, not ready yet

As already discussed previously~\cite{catte16}, NaCl has a negligible effect 
on the headgroup order parameters with submolar 
concentrations due to the very weak binding in PC bilayers. 
In contrast to almost all the other simulation models, this behaviour 
is reproduced by Lipid14 model with \AA{}qvist ions \cite{catte16}. 
The headgroup order parameter response to CaCl$_2$ concentration was,
however, overestimated with the same combination of force field
parameters~\cite{catte16}. The results were not improved when
ion models by Dang et al.~\cite{smith94,chang1999,dang2006} or
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
with more realistic bulk behaviour were used, as seen in 
in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
Together with extensive data in previous work \cite{catte16}, 
the results suggest that CaCl$_2$ binding behaviour in PC bilayers
cannot be corrected by only improving the ion models.

Significant improvement is achieved when the developed ECC-lipid model 
is used. The headgroup order parameter changes as a function of CaCl$_2$
concentration model show a good agreement between ECC-lipid model and 
experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
Ion binding affinity is quantified by using the relative surface
excess between ions and water from Eq.~\ref{surfexcess} and  
density profiles along membrane normal shown in Fig.~\ref{fig:cacl-dens}.
%This is directly related to the amount of surface excess charge 
%as shown along with ion density profiles in  Fig.~\ref{fig:cacl-dens}.
The relative surface excess for ECC-lipid model with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$
is significantly smaller than for Lipid14 with \AA{}qvist $\Gamma_i^w = XX nm^{-3}$
and Dang ions $\Gamma_i^w = 0.3 nm^{-3}$.
%The density profiles from simulations with original Lipid14 \cite{dickson14}
%and Dang ions \cite{smith94,chang1999,dang2006} 
%show a pronounced peak in the position of the phosphate moieties of POPC. 
%The use of ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
%or the ion model by \AA{qvist}~\cite{aqvist90}
%along with Lipid14 does not significantly change it 
%(Headgroup order parameter responses for these models in SI).
%The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
%exhibits on the other hand smaller density in this region 
%suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
%It is then likely that the models that overestimate head group order parameter response~\cite{catte16}
%also overestimate the amount of bound cations.
\todo{This paragraph is to be finalized when we have the surface excess analysis
  done by using Eq.~\ref{surfexcess}.}

@@

>>>>

Experimentally NaCl has a negligible effect on the headgroup order parameters
with submolar concentrations due to the very weak binding of Na$^+$
in PC bilayers \cite{akutsu81,catte16}. 
While almost all the available simulation models overestimate cation
binding in PC lipids, the weak binding of Sodium ions was reproduced by
Lipid14 model when simulated with \AA{}qvist ions~\cite{catte16}.
However, the headgroup order parameter response to CaCl$_2$ concentration was
overestimated with the same combination of force field parameters~\cite{catte16}.
Changing to ion models by Dang et al.~\cite{smith94,chang1999,dang2006} or
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
with more realistic bulk behaviour were used did not improve the results
for the CaCl$_2$ interactions with Lipid14 model, as seen in 
in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
Together with extensive data in previous work \cite{catte16}, 
the results suggest that CaCl$_2$ binding behaviour in PC bilayers
cannot be corrected by only improving the ion models.

Significant improvement is achieved when the developed ECC-lipid model 
is used together with ECC-ions. The headgroup order parameter changes
as a function of CaCl$_2$ concentration show a good agreement between
ECC-lipid model and experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
As discussed in previous section, also the response of headgroup order
parameters to the bound charge is in good agreement with the experiments in this
model. Thus, we conclude that the ECC-lipid model with ECC-ion parameters
gives a realistic binding affinity for the Na$^+$ and Ca$^{2+}$ binding
in POPC bilayer.

Based on the ion density profiles along membrane normal
in Fig.~\ref{fig:cacl-dens}, the Ca$^{2+}$ binding seems to
be stronger in Lipid14 model with Dang and ECC-ions.
This is in agreement 
This is quantified by calculating the relative surface
excess from Eq.~\ref{surfexcess}. This gives 
$\Gamma_i^w = 0.07 nm^{-3}$ for ECC-lipid model simulated with ECC-ions,
which is significantly smaller than $\Gamma_i^w = XX nm^{-3}$ for
Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 nm^{-3}$ with Dang ions.
\todo{Surface excess for Lipid14 with Aqvist yet to be added.} \\
\todo{Surface excess units are not correct. Maybe the numbers needs to be checked also?} \\
Rough estimates for the free energy difference between bound and unbound cation
can made from
\begin{equation}
  \Delta G={\rm k_b T}\log(\frac{p_o}{p_i}),
\end{equation}
where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
in the maximum density in bilayer, respectively. The density profiles in
Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
in Lipid14 with \AA{}qvist.

@@

Since ECC--lipid model gives a good agreement with experiments for the response
of headgroup order parameters to the bound charge as well as to the CaCl$_2$
concentration,
